Inference for Trans-dimensional Bayesian Models with Diffusive Nested Sampling

### 2014-08-28 Documents/Paper/ms.tex
citations

In this paper, we introduce an approach that uses trans-dimensional MCMC moves
to infer $N$, but embeds this process in the Nested Sampling framework to
overcome difficulties that would be encountered by sampling the posterior
distribution directly. Our motivation
for using Nested Sampling is not primarily to calculate the marginal
likelihood $\mathcal{Z}$ (including summing over $N$, as in
Equation~\ref{eq:evidence}),
although $\mathcal{Z}$ is readily available from the output.
The motivation for using Nested Sampling, rather than just sampling the
posterior using trans-dimensional MCMC, is that the posterior often contains
difficult features, such as strong dependencies or multiple modes, that cause
problems with mixing. Diffusive Nested Sampling
(DNS) replaces the posterior distribution with an alternative target
distribution composed of a mixture of the prior distribution with other, more
constrained versions of the prior, facilitating mixing between the multiple modes
and along degeneracy curves.
This is similar to the idea of annealed importance sampling
\citep{neal} or parallel tempering \citep{pt},
where the target distribution is modified from the posterior to something
``easier''. However,
the sequence of distributions used in Nested Sampling avoids some of the
problems with annealing-based methods. In particular, there is no need to choose
a temperature schedule (a potentially large set of tuning parameters),
and the method does not fail when phase transitions
are present.

@@

\section{Sinusoidal example}\label{sec:sinewaves}
In this section we demonstrate a seemingly simple example which exhibits
a phase transition, making a standard MCMC approach difficult. Suppose
a signal is composed of $N$ sinusoids, of different periods, amplitudes,
and phases. The signal is observed at various times $\{t_i\}_{i=1}^N$ with
noise, and we want to use the resulting data to infer the number of sinusoids
$N$, along with the periods $\{T_i\}_{i=1}^N$, amplitudes $\{A_i\}_{i=1}^N$,
and phases $\{\phi_i\}_{i=1}^N$.
This kind of model has many applications, and has been solved analytically
under certain sets of assumptions \citep[see e.g.][]{bretthorst}. Similar
models have been used in many different fields
\citep[e.g.][]{2005PhRvD..72b2001U, 2007ApJ...654..551B, 2003AIPC..659....3B}.
Note that this problem is also very closely related to the
problem of detecting exoplanet signals in radial velocity data, which has
attracted a lot of research attention in recent years
\citep[e.g.][]{2007ApJ...654..551B, gregory, fengji, 2011MNRAS.415.3462F}.

@@

>>>>

In this paper, we introduce an approach that uses trans-dimensional MCMC moves
to infer $N$, but embeds this process in the Nested Sampling framework to
overcome difficulties that would be encountered by sampling the posterior
distribution directly. Our motivation
for using Nested Sampling is not primarily to calculate the marginal
likelihood $\mathcal{Z}$ (including summing over $N$, as in
Equation~\ref{eq:evidence}),
although $\mathcal{Z}$ is readily available from the output.
The motivation for using Nested Sampling, rather than just sampling the
posterior using trans-dimensional MCMC, is that the posterior often contains
difficult features, such as strong dependencies or multiple modes, that cause
problems with mixing. Diffusive Nested Sampling
\citep[DNS][]{dnest} replaces the posterior distribution with an alternative target
distribution composed of a mixture of the prior distribution with other, more
constrained versions of the prior, facilitating mixing between the multiple modes
and along degeneracy curves.
This is similar to the idea of annealed importance sampling
\citep{neal} or parallel tempering \citep{pt},
where the target distribution is modified from the posterior to something
``easier''. However,
the sequence of distributions used in Nested Sampling avoids some of the
problems with annealing-based methods. In particular, there is no need to choose
a temperature schedule (a potentially large set of tuning parameters),
and the method does not fail when phase transitions
are present.

@@

\section{Sinusoidal example}\label{sec:sinewaves}
In this section we demonstrate a seemingly simple example which exhibits
a phase transition, making a standard MCMC approach difficult. Suppose
a signal is composed of $N$ sinusoids, of different periods, amplitudes,
and phases. The signal is observed at various times $\{t_i\}_{i=1}^N$ with
noise, and we want to use the resulting data to infer the number of sinusoids
$N$, along with the periods $\{T_i\}_{i=1}^N$, amplitudes $\{A_i\}_{i=1}^N$,
and phases $\{\phi_i\}_{i=1}^N$.
This kind of model has many applications, and has been solved analytically
under certain sets of assumptions \citep[see e.g.][]{bretthorst}. Similar
models have been used in many different fields
\citep[e.g.][]{2003AIPC..659....3B, 2005PhRvD..72b2001U, 2007ApJ...654..551B,
2009MNRAS.395.2226B}.
Note that this problem is also very closely related to the
problem of detecting exoplanet signals in radial velocity data, which has
attracted a lot of research attention in recent years
\citep[e.g.][]{2007ApJ...654..551B, gregory, fengji, 2011MNRAS.415.3462F}.

@@

\bibitem[Brewer and Stello(2009)]{2009MNRAS.395.2226B} Brewer, B.~J., 
Stello, D.\ 2009.\ Gaussian process modelling of asteroseismic data.\ 
Monthly Notices of the Royal Astronomical Society 395, 2226-2233.
